---
title: "R Data Types"
author: "Alex Vand"
date: "8/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vector 1
```{r}
vec_1 <- c(2, 5, 9, 10, 8, 12, 1, 0)
vec_1
```

Check the following for that vector:

- What is the class of the vector? class()
- What type of variable does it store? typeof()
- Access the 3rd element and store as vec_1_e3
- Access the 1st element and store as vec_1_e1
- Access the 5th through 7th elements and store as vec_1_e5to7
- Reassign vec_1 as a character using as.character, stored as vec_1_char. What does the output look like?

```{r}
class(vec_1)
typeof(vec_1)
vec_1_e3 <- vec_1[3]
vec_1_e3
vec_1_e1 <- vec_1[1]
vec_1_e1
vec_1_e5to7 <- vec_1[5:7]
vec_1_e5to7
vec_1_char <- as.character(vec_1)
vec_1_char
```

## Vector 2

```{r}
vec_2 <- c(town = "Santa Barbara", location = "Rincon", swell = "south")
vec_2
class(vec_2)
typeof(vec_2)
vec_2[2]
vec_2["location"]
```
- What is the class of vector elements? class()
- What is the length of vec_2?
- Access the 2nd element by name and store as vec_2_e2

```{r}
class(vec_2)
length(vec_2)
vec_2_e2 <- vec_2[2]
vec_2_e2
```

## Create a data frame in R, index

```{r}
df_1 <- data.frame(region = c("A", "B", "A", "D"),
                   species = c("otter", "great white", "sea lion", "gray whale"),
                   count = c(12, 2, 36, 6))
df_1
```

- Return the class of the entire data frame
- Return the class of the species column
- Find the maximum value of the count() column, store as max_count

```{r}
class(df_1)
class(df_1$species)
class(df_1[[3]])
max_count <- max(df_1$count)
max_count
```




