---
title: "Salamander & Trout Data"
author: "Alex Vand"
date: "8/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

## Read in the data

```{r}
mack_verts <- read_csv(here::here("data", "AS00601.csv"))
mack_verts

# Update the names to lower_snake_case

mack_verts <- clean_names(mack_verts)
names(mack_verts)
```

## Data wrangling and exploration

```{r}
# Store the 5th value in column “WEIGHT” as `mc_wt_5`. Check by looking at your data frame to confirm.
mc_wt_5 <- mack_verts$weight[5]
mc_wt_5

# Store the 8th - 20th value in the “LENGTH1” column as `mc_length_8_20`. Check by looking at your data frame to confirm.
mc_length_8_20 <- mack_verts$length1[8:20]
mc_length_8_20

# Store everything in column SAMPLEDATE as a vector called `mc_dates`.
mc_dates <- mack_verts$sampledate
length(mc_dates) # to see number of entries in this vector
```

## Make a salamander subset
Create a subset that only contains observations for Pacific Giant Salamanders (species _Dicamptodon tenebrosus_ , stored in `species` as 'DITE'). Store the subset as `mc_salamanders`. _Hint_ : see `dplyr::filter()`!

```{r}
mc_salamanders <- filter(mack_verts, species == 'DITE')
mc_salamanders
length(mc_salamanders) # number of columns
length(mc_salamanders$species) # number of rows
```


## Make a scatterplot

- Create a scatterplot of length1 (snout-vent length in millimeters) versus weight (grams) for all salamanders in the dataset. Update axis labels, title, subtitle, and add a caption with the data source. Customize point color and size, possibly opacity, and theme.
- Export your scatterplot as `salamander_size.png` to your figs folder.

```{r}
salamander_size <-
  ggplot(data = mc_salamanders, aes(x = length1, y = weight)) +
  geom_point(color = "orange", alpha = 0.5, shape = "square") +
  labs(title = "Salamander Length (mm) vs. Weight (g)",
       subtitle = "snout-vent length for Dicamptodon species",
       caption = "Data source: EDI Data Portal", 
       x = "Length (mm)",
       y = "Weight (g)") +
  theme_grey()

ggsave(here::here("figs", "salamander_size.png"))
```


## Make a cutthroat plot

- Similar to above, make a subset called mc_trout that only contains observations for cutthroat trout
- Create a scatterplot of length1 by weight for all trout in the dataset
- Customize so that the point color depends on reach
- Customize your color scheme (e.g. scale_color_manual())
- Facet your plot by creek reach (facet_wrap(~...))
- Update graph axis labels and title
- Export your graph as cutthroat_size.png to figs

```{r}
mc_trout <- filter(mack_verts, species == "ONCL")

cutthroat_size <-
  ggplot(data = mc_trout, aes(x = length1, y = weight)) +
  geom_point(aes(color = reach)) +
  scale_color_manual(values = c("red", "blue", "yellow")) +
  facet_wrap(~reach) +
  labs(x = "Length (mm)",
       y = "Weight (g)",
       title = "Trout Length (mm) vs. Weigth (g)")

ggsave(here::here("figs", "cutthroat_size.png"))
```






